---
title: Traffic Data Analysis
output: html_document
---

##initialization
```{r}
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)
library(readr)
library(corrplot)

data.raw <- read.csv("C:/Users/USER/Desktop/csen1061-project1-traffic-data-analysis/all-semi-unique.csv")
```
First we take a look at the data
```{r}
glimpse(data.raw)
```

## removing ads
We found some ads in the data, they won't be of any benefit to us so we have to remove them
```{r}
data.road <- data.raw %>% select(-(starts_with("ad.")))
```

## removing images
We also found some images that do not represent any thing so we will delete them as well
```{r}

data.road <- data.road %>% select(-(rd.rp.rpImg), -(rd.rp.img))

data.road <- data.road %>% select(-(rd.img))
```

## removing constant columns
I saw that there are some columns of the same value so after testing i found that there are 2 columns rd.cl and rd.rp.type with the same value for all unput, these also won't be of any benefit and should be removed.
```{r}
data.road %>%sapply(unique) %>%sapply(length)

data.road <- data.road %>% select(-(rd.cl), -(rd.rp.type))
```
## formating the date
we saw that the date has a format that isn't really useful so we have to reformat it.
```{r}
head(data.road$crawl_date)

data.road$crawl_date <- strptime(data.road$crawl_date, format = "%a %b %e %X UTC %Y", tz = "UTC") %>% as.POSIXct()

head(data.road$crawl_date)
```

## estimating the actual report time
```{r}
data.road <- data.road %>% mutate(report_time = as.POSIXct(round(crawl_date - (rd.rp.hr*60*60 + rd.rp.mn*60), "mins"))) %>% select(-c(rd.rp.mn, rd.rp.hr))
```

## dividing the road into a primary and secondary streets
i found that some of rd.nm column has 2 values separated by a ; so lets get them actually separated.
```{r}
glimpse(data.road)

data.road <- data.road %>% separate(rd.nm, c("rd.primary", "rd.secondary"), ";")
```
and some secondary roads weren't present so they introduced some new NA's and we should reconstruct them to be able to come up with any data later.
```{r}
data.road$rd.secondary[is.na(data.road$rd.secondary)] <- "general"

glimpse(data.road)
```
## getting status reports
we can separate the status of the reports of the data for further investigation.
```{r}
data.road.status <- data.road %>% select(crawl_date, rd.ri, rd.stid) %>% unique()

glimpse(data.road.status)
```

## removing duplicates
we found some duplicates in the data because of the way the data were crawled so we have to remove it.
```{r}
data.road <- data.road[!(data.road$rd.rp.cmid %>% duplicated),]
```
## seperating all reports alone
lets take a deeper look at the reports alone by separating them.
```{r}
data.road.reports <- data.road %>%  select(rd.ri, rd.rp.fullnm, rd.rp.stid, rd.rp.cm, rd.rp.cmid, report_time, rd.primary, rd.secondary) %>% unique()
```
## trying to get crowded info
trying to summarize the data to know which roads are the most crowded.
```{r}
crowded_roads <- data.road.reports %>% group_by(rd.primary) %>% summarise(crowded = mean(rd.rp.stid))
```
